Writing Trash Bin

### 2024-03-03 paper.qmd
update introduction

## Introduction

> Edit note: this section is untouched since the confirmation. Might need a rewrite. should comment that visualisation helps us understand *how* our data is different.

Think back to the last time you made some sort of data visualisation. What was the purpose of that visualisation? Was it to better understand your data? Was it to help you make a decision? Was it to to communicate that decision to someone else? Now think about the last time you expressed some form of uncertainty. Was it a set of numerical confidence intervals? Maybe they were expressed as a set of values in a table. Did you consider visualising your uncertainty instead? There are many stages in our analysis that benefit from the power of data visualisation, however this does not mean it is always done with success. Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data driven discovery is highlighted by data sets such as Anscombe's quartet [@anscombe] or the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information when visualised. Anscombe quartet is shown in @fig-anscombe, because describing the data is never the same as seeing it. Instead of having to repeatedly check endless hypothesis to find interesting numerical features, visualisations **tell** us what is important about our data. This powerful aspect of data visualisation is poorly or seldom used in later stages when we are communicating our findings, specifically with respect to uncertainty.

@@

## Similar work
> Edit note: I think with the addition of more summary papers I should just summarise common themes of issues the papers ran into instead of detailling the work paper by paper.

There are a wealth of uncertainty communication and visualisation literature reviews that already exist. Interestingly they all seem to have identified similar pitfalls or fallen into one themselves.

@@

>>>>

## Introduction
From entertainment choices to news articles to insurance plans, the modern citizen is so over run with information in every aspect of their life it can be overwhelming. In this overflow of information, tools that can effectively summarise information down into simple and clear ideas become more valuable. Information visualisations remain one of the most powerful tools for fast and reliable science communication.

There are many stages in our analysis that benefit from the power of data visualisation, however this does not mean it is always done with success. Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data driven discovery is highlighted by data sets such as Anscombe's quartet [@anscombe] or the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information when visualised. Anscombe quartet is shown in @fig-anscombe, because describing the data is never the same as seeing it. Instead of having to repeatedly check endless hypothesis to find interesting numerical features, visualisations **tell** us what is important about our data and **how** it might diverge from what we expect.

@@

Uncertainty visualisation is a somewhat new subfield of visualisation, with early papers that specifically reference an "uncertainty visualisation" popping up in the late 90s and early 2000s (*Cite Gap 1*). Despite the youth of the field, commonly used uncertainty visualisations such as a box plot or histogram have been around for almost as long as statistical graphics themselves (*Cite Gap 2*). Early mentions of "uncertainty visualisation" appear in computer science papers discussing uncertainty visualisation as a subfield of geospatial information visualisation, specifically for the many uncertainties that come with satelite images (*Cite Gap 1*). In the modern day, uncertainty visualisation has exploded into its own field that applies to a wide range of contexts and is no longer bound by geospatial information (*Cite Gap 1*).

## Similar work
::: aside
I have a large backlog of uncertainty literature review summaries in my Mendely that need to be added here. That is very high on my todo list once the readable draft is done.
:::

Any field that experiences an explosion of new research will also invite a series of literature reviews that seek to combine and summarise that information. Given that uncertainty visualisation has been around for a few decades now, there is already a wealth of uncertainty visualisation literature reivews. Interestingly these papers seem to have identified a series of common pitfalls in uncertainty visualisation and its evaluation.

@@

> Edit note: just summarise common themes of issues the papers ran into instead of detailling the work paper by paper.

### 2024-03-03 writingdumpster.qmd
update introduction

@@

# Previous literature review
##### Introduction addition
From entertainment choices to news articles to insurance plans, the modern citizen is so over run with information in every aspect of their life it can be overwhelming. So overwhelming in fact, that it is not uncommon for people to become so overwhelmed by the information they refuse to engage with it at all. In this overflow of information, tools that can effectively summarise information down into simple and clear ideas become more valuable. While websites that summarise insurance plans or hotel options down into a clear table make a living off it, few tools are as powerful or effective as a visualisation. Information visualisations remain one of the most powerful tools for fast and reliable science communication. Effectively leveraging visualisations to make them more effective is therefore key in facilitating simple and clear communication of science and other ideas.

>>>>

# Citation Gaps
Cite Gap 1: When did uncertainty visualisaiton become its own subfield of visualisation?

Cite Gap 2: When was the first "uncertainty visualisation" used (likely a box plot of histogram)

@@

# Previous literature review
